---
title: "Migration Data Visualisation"
author: "Erik"
date: '2018-03-10'
slug: immigration-data
tags: []
categories: []
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/erik/Documents/blog/data")
```

In this post I will be visualizing the migration pattern of the US with data from 2012 using a chord diagram from the chorddiag package which in turn is using the JavaScript visualization library D3.
The data I'm going to use are stored in two files.  

```{r, message=F, warning=F}
library(readr)
library(tidyverse)
dat <- read_csv("migration2016.csv")
states <- read_csv("states_chord.csv")
head(dat[1:5],5)
```

We can for example see that 423 people moved from Alabama to Alaska. The `states` data set contains regions, states and colors that I plan to use for the chord diagram. 

To prepare the data for the package, I need a matrix and some further data manipulation to get the data in the right order.

```{r, message=F}
library(reshape)
dat <- melt(as.data.frame(dat),id.vars = c("To"))
colnames(dat) <- c("To","From","value")
merged = merge(x=dat, y=states, by.x="From", by.y="States")
colnames(merged)[4] = "ID_From"
merged = merge(merged, states, by.x = "To", by.y="States")
colnames(merged)[7] <- "ID_To"
merged <- merged %>% arrange(ID_From) %>% arrange(ID_To)
done <- matrix(merged$value,byrow = T, nrow=51)
colnames(done) <- unique(merged$To) ; rownames(done) <- unique(merged$To)
```

Let's take a look at the top left corner of the 52 times 52 matrix

```{r}
head(done[,1:5],5)
```


And finally the diagram is created. Hovering a specific 
```{r}
library(chorddiag)
groupNames = unique(merged$To)
colnames(done) <- groupNames
row.names(done) <- groupNames
groupColors <- merged$Color.x
done[is.na(done)]<-0

chorddiag(done,groupColors = groupColors,
          showTicks = F,
          width = 700,
          height = 700)


```


