---
title: Swedish politics and Twitter
author: Erik
date: '2018-04-05'
slug: swedish-politics-and-twitter
categories: []
tags: []
---
```{r, include=F}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(eval = T)
knitr::opts_knit$set(root.dir = "/Users/erik/Documents/blog/data")
```

In this post I'll be using the package 'rtweet' to look at how the Swedish political parties have used twitter in the recent years. I will also show how to inspect individual twitter accounts, for example the party leaders, to see what words they frequently use on twitter.

The package is very easy to use for downloading tweets, both from invidiual accounts and hashtags. For example, I could download the last, say 100 tweets, tagged with the popular hashtag #svpol by running

```{r}
rt <- rtweet::search_tweets("#svpol", n = 100, include_rts = FALSE)
```

Let's look at the last five tweets and see what we can figure out about them. For example, we can see their twitter names, the actual tweets and profile picture URL

```{r}
head(rt$screen_name,3)
head(rt$text,3)
head(rt$profile_image_url,3)
```


There's actually a whole lot of things we can find out about these tweets. All in all around 80 variables. Here are the 20 first

```{r}
head(colnames(rt),20)
```

## Plotting activity over time

Let's begin with looking at the twitter activity. I will be using ggplot and plotly. I first create a function that downloads all the parties into a list of datasets. I then rbind them, fix labels and colours so that they correspond to the traditional colors used for the parties, fix the dates and plot.
```{r}
# library
library(rtweet)
library(dplyr)
library(lubridate)
library(ggplot2)
library(plotly)

# load tweets
imp_tweets <- function(n, include_rts=T){
  parti <- list()
  parti[["v"]] <- get_timeline("vansterpartiet",n, include_rts=include_rts)
  parti[["s"]] <- get_timeline("socialdemokrat",n, include_rts=include_rts)
  parti[["mp"]] <- get_timeline("miljopartiet", n, include_rts=include_rts)
  parti[["sd"]] <- get_timeline("sdriks",n, include_rts=include_rts)
  parti[["c"]] <- get_timeline("Centerpartiet",n, include_rts=include_rts)
  parti[["l"]] <- get_timeline("liberalerna",n, include_rts=include_rts)
  parti[["m"]] <- get_timeline("nya_moderaterna",n, include_rts=include_rts)
  parti[["kd"]] <- get_timeline("kdriks",n, include_rts=include_rts)
  return(parti)
}

# download 2000 tweets and includ retweets
parti <- imp_tweets(2000, T)

# create a df with all the dates
act <- lapply(parti, function(x) x[,c("created_at","screen_name")])
tweets <- do.call(rbind.data.frame, act)

# plot activity if tweets is a list of data frames of different accounts
plotpartytwitter <- function(floor_to = "day"){
  tweets$screen_name <- as.factor(tweets$screen_name)
  labels = c("Centerpartiet","Kristdemokraterna","Liberalerna","Miljöpartiet","Nya Moderaterna","Sverigedemokraterna","Socialdemokraterna","Vänsterpartiet")
  values = c("#15983B", "#231D78", "#6FB9EA","#81CF44","#204FD9","#DDDB30","#EB242B","#97040C")
  levels(tweets$screen_name) <- labels
  
  tweets_act <- tweets %>%
    #mutate(created_at = format.POSIXct(created_at, tz = "Europe/Stockholm")) %>% 
    mutate(created_at = format(floor_date(as.POSIXct(created_at), floor_to))) %>% 
    mutate(time = as.POSIXct(created_at)) %>% 
    group_by(screen_name) %>% 
    count(time) 
  
  ftt <- floor_to  %>% as.data.frame() %>% mutate(floor_to = ifelse(floor_to=="week","vecka",
                                              ifelse(floor_to=="month","månad",
                                                     ifelse(floor_to=="hour","timma",NA))))
  
  g = ggplot(tweets_act, aes(x=time,
                             y=n,
                             color=screen_name,
                             text=paste0(screen_name, "<br>Antal tweets denna ",ftt$floor_to,": ",n)))+
    geom_line()+
    theme_minimal(base_size = 14)+
    labs(title = "Partiernas aktivitet på Twitter", x = paste0("Antal tweets (per ", floor_to ,")" ,sep="" ) , y = "", color ="Parti\n")+
    scale_color_manual(values=values)
  ggplotly(g, tooltip=c("text")) %>% config(displayModeBar = F)
}
```


Now all that remains is to run the function. It is possible to hide individual parties by clicking at them, zoom in on different times periods. Double click to zoom out again. I can change between month, day, hour, minute if I want. For this plot, I will look at month. This flexibility is great because I can easily turn this into a Shiny app.
```{r}
plotpartytwitter("month")
```


## Plotting popular words with wordcloud

The package wordcloud makes creating beautiful wordclouds incredibly easy. All one has to do is fix the data, remove some uninteresting words, get rid of some symbols and emojis and you're ready to go. Here, I will look at the last 1500 tweets by Annie Lööf. 


```{r}
# library
library(wordcloud)
library(tm)
library(readr)
library(RColorBrewer)
library(ggthemes)


# download Last 1500 tweets from Annie Lööf without retweets
tweets_no_rt <- get_timeline("annieloof", 1500, include_rts=F)

# get rid of smileys , curly quotes, dash and dots
twtext <- tweets_no_rt$text %>% lapply(function(x) iconv(x,"UTF-8-MAC","UTF-8", sub=""))
twtext <- gsub("[“”]", "", twtext)
twtext <- gsub('[\u2013:\u2016]', "", twtext)
twtext <- gsub("…","", twtext)

# create corpus
tw_c <- Corpus(VectorSource(twtext))

# clean data from stopwords, punctuations, numbers and make everything lower case
stopwords_sv <- read_csv("stopwords-sv.txt")

tw_c <- tw_c %>%
  tm_map(removePunctuation) %>% 
  tm_map(content_transformer(tolower)) %>% 
  tm_map(removeNumbers) %>%
  tm_map(stripWhitespace) %>% 
  tm_map(removeWords, stopwords("sv")) %>%
  tm_map(removeWords, as.matrix(stopwords_sv))


# plot wordcloud
pal <- brewer.pal(9,"Dark2")
pal <- pal[-(1:4)]
wordcloud(tw_c,max.words = 150,random.order = F,colors = pal,scale = c(4,0.1))
```

To create a more traditional plot to see frequency of each of the words. Hovering over each bar will display exact number of times the word was used

```{r}
# plot most frequent used words
tdm <- TermDocumentMatrix(tw_c)
tdm.rs <- sort(rowSums(as.matrix(tdm)), decreasing=TRUE)
tdm.df <- data.frame(word = names(tdm.rs), freq = tdm.rs, stringsAsFactors = FALSE)

ggplotly(ggplot(head(tdm.df,20), aes(x=reorder(word,freq), y=freq, text=paste0("Antal gånger ordet ",word," användes: ",freq)))+
  geom_bar(stat="identity")+
  coord_flip()+
  ylab("Frequency")+
  xlab("")+
  theme_hc(), tooltip=c("text"))%>% config(displayModeBar = F)
```



