---
title: "Shiny App: Airline Routes"
author: "Erik"
date: '2018-03-08'
slug: flights
tags: []
categories: []
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/erik/Documents/blog/data")
#devtools::install_github("bwlewis/rthreejs")
```

In this post I am going to create a simple Shiny app that will show different airlines and their routes. First of all I will pick three airlines and create a plot, and from there incoroprate it into a Shiny app with a few inputs. The data [are](https://www.theguardian.com/news/datablog/2010/jul/16/data-plural-singular, "Data is or data are?") from openflights.org. 

```{r}
airports <- read.csv("airports.dat", header=FALSE, stringsAsFactors=FALSE)
colnames(airports) <- c("airport_id", "name", "city", "country", "IATA_FAA", "ICAO", "latitude", "longitude", "altitude", "timezone", "DST")
airports <- subset(airports, IATA_FAA != '')
flights <- read.csv("routes.dat", header=FALSE, stringsAsFactors=FALSE)
colnames(flights) <- c("airline", "airline_id", "source_air", "source_air_id", "dest_air", "dest_air_id","code_share", "stops", "equipment")
```

For the first plot, I select three airlines. EK: Emirates, LH: Lufthansa, DL: Delta

```{r}

suppressMessages(library(dplyr))
flights <- flights %>% 
  filter(airline %in% c("EK","LH","DL"))

```


Prepare data for the `globejs()` function
```{r}
suppressMessages(library(threejs))
f1 <- flights %>% select(airline, source_air, dest_air)
a1 <- airports %>% select(IATA_FAA, latitude, longitude )

m <- merge(x = f1, y=a1, by.x = "source_air", by.y = "IATA_FAA")
m <- merge(x = m,  y=a1, by.x = "dest_air", by.y="IATA_FAA")

m <- m %>% select(airline, latitude.x, longitude.x, latitude.y, longitude.y, airline)
colnames(m) <- c("airline","origin_lat","origin_long","dest_lat","dest_long")

m <- m %>% 
  arrange(airline) %>% 
  mutate(color=ifelse(airline=="DL","#02C5FE",
                      ifelse(airline=="LH","#DE88F78",
                             ifelse(airline=="EK","#88F8AB",NA))))


globejs(arcs=m[,2:5], long=airports[,8],lat=airports[,7],value=0.1,arcsHeight=0.23, arcsLwd=0.01, arcsColor=m$color, color="white",arcsOpacity=.25, emissive="black",
        atmosphere=TRUE, height=700, width = 700, bodycolor='white')

```

Now we can turn this into a shiny app with a few inputs, for example which airlines to use etc. This is a static page created with blogdown, and since shiny apps depend on a live R session, I've deployed the following app to shinyapps.io. Below is the code used to create the app and [HERE](https://erst.shinyapps.io/flights/) is a link to the app. I've also included the app in a frame on this page.

```{r, eval=F}
#devtools::install_github("bwlewis/rthreejs")
library(threejs)
library(shiny)
library(dplyr)
library(shinythemes)
library(shinydashboard)


# Define UI for application that draws a histogram
ui <- fluidPage(
  theme = shinytheme("cyborg"),
  

      # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
        
        tags$div(class = "header", checked = NA,
             tags$b("Select maximum three airlines using their IATA codes"),
             tags$br(),
             tags$em("i.e. Lufthansa: LH, Turkish: TK")
        ),
        
        uiOutput("airSelector"),
        sliderInput(inputId="height",label="Height:", min = 1, max = 5, value=1),
         sliderInput(inputId="width",label="Width:", min = 1, max = 5, value=1),
         sliderInput(inputId="opac",label="Opacity:", min = 0, max = 1, value=0.3),
         checkboxInput(inputId="atmos", label="Atmosphere",value = TRUE, width = "800px"),
         
         tags$div(class = "header", checked = NA,
                  tags$a(href = "http://www.sterlingtravel.co.uk/airline_list.html", "Airline names and corresponding IATA codes"))
         
         

      ),
      
      
      
      # Show a plot of the generated distribution
      mainPanel(
         globeOutput("flightPlot", width = "100%", height = "1000px")
      )
   )
)



# Define server logic required to draw a histogram
server <- function(input, output) {

  airports <- read.csv("./data/airports.dat", header=FALSE, stringsAsFactors=FALSE)
  colnames(airports) <- c("airport_id", "name", "city", "country", "IATA_FAA", "ICAO", "latitude", "longitude", "altitude", "timezone", "DST")
  airports <- subset(airports, IATA_FAA != '')
  flights <- read.csv("./data/routes.dat", header=FALSE, stringsAsFactors=FALSE)
  colnames(flights) <- c("airline", "airline_id", "source_air", "source_air_id", "dest_air", "dest_air_id","code_share", "stops", "equipment")
  
  ### for select options
  output$airSelector <- renderUI({
    selectizeInput(inputId = "selectAirline",
                   label = " ", 
                   multiple = T,
                   options = list(maxItems = 3, placeholder = 'Ex: LH SK TK'),
                   choices=rev(unique(flights$airline))) 
  })
  
  
   output$flightPlot <- renderGlobe({

     
     flights2 <- flights %>% 
       filter(airline %in% input$selectAirline)

     f1 <- flights2 %>% select(airline, source_air, dest_air)
     a1 <- airports %>% select(IATA_FAA, latitude, longitude )
     
     m <- merge(x = f1, y=a1, by.x = "source_air", by.y = "IATA_FAA")
     m <- merge(x = m,  y=a1, by.x = "dest_air", by.y="IATA_FAA")
     
     m <- m %>% select(airline, latitude.x, longitude.x, latitude.y, longitude.y, airline)
     colnames(m) <- c("airline","origin_lat","origin_long","dest_lat","dest_long")
     
     m <- m %>% 
       arrange(airline) %>% 
       mutate(color=ifelse(airline==input$selectAirline[1],"#02C5FE",
                           ifelse(airline==input$selectAirline[2],"#DE88F78",
                                  ifelse(airline==input$selectAirline[3],"#88F8AB",NA))))
     globejs(arcs=m[,2:5],
             long=airports[,8],
             lat=airports[,7],
             value=0.1,
             arcsHeight=input$height/10,
             arcsLwd=input$width/2,
             arcsColor=m$color,
             color="white",
             arcsOpacity=input$opac,
             emissive="black",
             atmosphere=input$atmos,
             bodycolor='white',
             fov=35, lightcolor="grey",
             rotationlong=42, rotationlat=57,
             bg = "#060606")

   })
}

# Run the application 
shinyApp(ui = ui, server = server)




```


<iframe src="https://erst.shinyapps.io/flights/", width="1000" height="800"></iframe>
